###Inferential Data Analysis

```{r include=T, eval=T, echo=F}
    suppressWarnings(library(plyr));
    suppressWarnings(library(ggplot2));
```
  
  
####1. Load the ToothGrowth data and perform some basic exploratory data analyses.
  
**Fig 1.1**  The boxplot graph below appears to indicate that the supplement in OJ results in higher average tooth growth than the supplement in VC.  However, the amount of overlap between the two seems to indicate that the difference may not be statistically significant.  The two supplements could on average perform the same.
```{r Exploratory1, include=T, eval=T, echo=T}
#   Load ToothGrowth Dataset    
    library(datasets);

#   The Effect of Dose on Tooth Growth
    g1 <- ggplot(ToothGrowth, aes(x=factor(supp), y=len, fill=factor(supp)))
    g1 <- g1 + geom_boxplot() 
    g1 <- g1 + ggtitle("Fig 1.1: Tooth Growth by Supplement") + labs(x="Supplement", 
        y="Tooth Length")
    g1
```

**Fig 1.2**  The boxplot graph below appears to indicate that average tooth growth and dose are positively correlated; as dose increases, tooth length also increases.  The lack of overlap between the three doses seems to indicate that the difference in tooth growth may be statistically significant.  
```{r Exploratory2, include=T, eval=T, echo=T}
#   The Effect of Dose on Tooth Growth
    g2 <- ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=factor(dose)))
    g2 <- g2 + geom_boxplot() 
    g2 <- g2 + ggtitle("Fig 1.2: Tooth Growth by Dose") + labs(x="Dose", y="Tooth Length")
    g2
```
  
**Fig 1.3**  The boxplot graph below appears to indicate that average tooth growth and dose are positively correlated for both supplements; as dose increases, tooth length also increases.  However, the supplement in OJ results in higher levels of average tooth growth for the 0.5mg and 1.0mg doses only.  There appears to be little or no difference between the supplments in OJ and VC for the 2.0mg dose.  
```{r Exploratory3, include=T, eval=T, echo=T}
#   The Combined Effects of Supplement & Dose on Tooth Growth
    g3 <- ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=factor(dose)))
    g3 <- g3 + geom_boxplot() + facet_grid(.~supp) 
    g3 <- g3 + ggtitle("Fig 1.3: Tooth Growth by Supplement & Dose") + labs(x="Dose", 
        y="Tooth Length")
    g3
```
  
####2. Provide a basic summary of the data.

The dataset describes the relationship between tooth length and various doses of vitamin C and absorbic acid (i.e., 0.5mg, 1.0mg, 2.0mg).  Summary of the ToothGrowth dataset is as follows:
```{r Summary, include=T, eval=T, echo=T}
    dim(ToothGrowth);
    head(ToothGrowth, n=5);
    summary(ToothGrowth);
```

####3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering).
  
This section compares tooth growth by supp and dose in terms of the following three scenarios: 1) tooth growth by supplement, 2) tooth growth by dose, and 3) tooth growth by supplement and dose.  
  
Question #1: Is there a difference in levels of tooth growth between supplements?  
**There is no statistically significant difference in the level of tooth growth between supplements.  See the box plot in Fig 1.1 above and the t-test results below.**  
  
T-Test #1  
H0: **There is no difference in tooth growth given the supplement by itself.**  
HA: The supplement by itself results in different levels of tooth growth.  
Results: p-value = 0.06063 --> **Fail to Reject the Null**
```{r Test1, include=T, eval=T, echo=T}
#   The Effect of the Supplement on Tooth Growth
    t.test(ToothGrowth$len[ToothGrowth$supp=="OJ"], ToothGrowth$len[ToothGrowth$supp=="VC"],
           con.levle=0.95, paired=F);
```
  
Question #2: Is there a difference in the level of tooth growth between doses? 
**There is a statistically significant difference in the level of tooth growth between doses.  See the box plot in Fig 1.2 above and the t-test results below.**  
  
T-Test #1  
H0: There is no difference in tooth growth given the amount of the dose.  
HA: **Larger doses result in increased levels of tooth growth.**
Results: p-value = 1.268e-07 --> **Reject the Null**  
```{r Q2T1, include=T, eval=T, echo=T}
    t.test(ToothGrowth$len[ToothGrowth$dose==1.0], ToothGrowth$len[ToothGrowth$dose==0.5],
           con.levle=0.95, paired=F);
```
  
T-Test #2  
H0: There is no difference in tooth growth given the amount of the dose.  
HA: **Larger doses result in increased levels of tooth growth.**
Results: p-value = 4.398e-14 --> **Reject the Null**  
```{r Q2T2, include=T, eval=T, echo=T}
    t.test(ToothGrowth$len[ToothGrowth$dose==2.0], ToothGrowth$len[ToothGrowth$dose==0.5],
           con.levle=0.95, paired=F);
```
  
T-Test #3  
H0: There is no difference in tooth growth given the amount of the dose.  
HA: **Larger doses result in increased levels of tooth growth.**
Results: p-value = 1.906e-05 --> **Reject the Null**  
```{r Q2T3, include=T, eval=T, echo=T}
    t.test(ToothGrowth$len[ToothGrowth$dose==2.0], ToothGrowth$len[ToothGrowth$dose==1.0],
           con.levle=0.95, paired=F);
```
  
Question #3: Is there a difference in level of tooth growth between supplements given the same dose?  
**There is a statistically significant difference in the level of tooth growth between supplements in OJ and VC for 0.5mg and 1.0mg does.  However, there is no statistically significant difference in the level of tooth growth between supplements in OJ and VC for 2.0mg doses.  See the box plot in Fig 1.3 above and the t-test results below.**   
  
T-Test #1  
H0: There is no difference in tooth growth given the supplement and the same dose.  
HA: **The supplement combined with the same dose results in increased levels of tooth growth.**  
Results: p-value = 0.006359 --> **Reject the Null** 
```{r Q3T1, include=T, eval=T, echo=T}
    t.test(ToothGrowth$len[ToothGrowth$supp=="OJ" & ToothGrowth$dose==0.5], 
        ToothGrowth$len[ToothGrowth$supp=="VC" & ToothGrowth$dose==0.5], 
        con.levle=0.95, paired=F);
```
  
T-Test #2  
H0: There is no difference in tooth growth given the supplement and the same dose.  
HA: **The supplement combined with same dose results in increased levels of tooth growth.**  
Results: p-value = 0.001038 --> **Reject the Null** 
```{r Q3T2, include=T, eval=T, echo=T}
    t.test(ToothGrowth$len[ToothGrowth$supp=="OJ" & ToothGrowth$dose==1.0], 
        ToothGrowth$len[ToothGrowth$supp=="VC" & ToothGrowth$dose==1.0], 
        con.levle=0.95, paired=F);
```
  
T-Test #3  
H0: **There is no difference in tooth growth given the supplement and the same dose.**  
HA: The supplement combined with same dose results in increased levels of tooth growth.  
Results: p-value = 0.9639 --> **Fail to Reject the Null**
```{r Q3T3, include=T, eval=T, echo=T}
    t.test(ToothGrowth$len[ToothGrowth$supp=="OJ" & ToothGrowth$dose==2.0], 
        ToothGrowth$len[ToothGrowth$supp=="VC" & ToothGrowth$dose==2.0], 
        con.levle=0.95, paired=F);
```
  
####4. State your conclusions and the assumptions needed for your conclusions. 
  

