###Inferential Data Analysis

```{r include=T, eval=T, echo=F}
    suppressWarnings(library(plyr));
    suppressWarnings(library(ggplot2));
```
  
  
####1. Load the ToothGrowth data and perform some basic exploratory data analyses.
  
The boxplot graph appears to indicate that the supplement in OJ results in higher average tooth lenth than the supplement in VC.
```{r Exploratory1, include=T, eval=T, echo=T}
    library(datasets);

#   The Effect of Dose on Tooth Growth
    g1 <- ggplot(ToothGrowth, aes(x=factor(supp), y=len, fill=factor(supp)))
    g1 <- g1 + geom_boxplot() 
    g1 <- g1 + ggtitle("Tooth Growth vs Supplement Only") + labs(x="Supplement", y="Tooth Length")
    g1
```

The boxplot graph appears to indicate that does and tooth length are positively correlated; as the dose increases, tooth length also increases.
```{r Exploratory2, include=T, eval=T, echo=T}
#   The Effect of Dose on Tooth Growth
    g2 <- ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=factor(dose)))
    g2 <- g2 + geom_boxplot() 
    g2 <- g2 + ggtitle("Tooth Growth vs Dose Only") + labs(x="Dose", y="Tooth Length")
    g2
```
  
The boxplot graph appears to indicate that does and tooth length are positively correlated for both supplements; as the dose increases, tooth length also increases.  It is interesting to note that OJ results in higher levels of tooth length for the 0.5mg and 1.0mg does.  However, there appears to be little or no difference for the 2.0mg dose.  
```{r Exploratory3, include=T, eval=T, echo=T}
#   The Combined Effects of Supplement & Dose on Tooth Growth
    g3 <- ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=factor(dose)))
    g3 <- g3 + geom_boxplot() + facet_grid(.~supp) 
    g3 <- g3 + ggtitle("Tooth Growth vs Dose by Supplement") + labs(x="Dose", y="Tooth Length")
    g3
```
  
####2. Provide a basic summary of the data.

The dataset describes the relationship between tooth length and various doses of vitamin C and absorbic acid (i.e., 0.5mg, 1.0mg, 2.0mg).  Summary of the ToothGrowth dataset is as follows:
```{r Summary, include=T, eval=T, echo=T}
    dim(ToothGrowth);
    head(ToothGrowth, n=5);
    summary(ToothGrowth);
```

####3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering).
  
Test #1: Does the supplement result in increased levels of tooth growth?  
H0: There is no difference in tooth growth given the supplement.  
HA: The supplement does result in different levels of tooth growth.  
```{r Test1, include=T, eval=T, echo=T}
#   The Effect of the Supplement on Tooth Growth
    t.test(ToothGrowth$len[ToothGrowth$supp=="OJ"], ToothGrowth$len[ToothGrowth$supp=="VC"],
           con.levle=0.95, paired=F);
```
  
Test #2: Does larger doses result in increased levels of tooth growth?  
H0: There is no difference in tooth growth given the amount of the dose.  
HA: Larger doses result in increased levels of tooth growth.  
```{r Test2, include=T, eval=T, echo=T}
#   The Effect of Dose on Tooth Growth
    t.test(ToothGrowth$len[ToothGrowth$dose==0.5], ToothGrowth$len[ToothGrowth$dose==1.0],
           con.levle=0.95, paired=F);

    t.test(ToothGrowth$len[ToothGrowth$dose==0.5], ToothGrowth$len[ToothGrowth$dose==2.0],
           con.levle=0.95, paired=F);

    t.test(ToothGrowth$len[ToothGrowth$dose==1.0], ToothGrowth$len[ToothGrowth$dose==2.0],
           con.levle=0.95, paired=F);
```
  
Test #3: Does combined effects of both the supplement and larger doses result in increased levels of tooth growth?  
H0: There is no difference in tooth growth given the supplement and the amount of the dose.  
HA: The supplement combined with larger does result in increased levels of tooth growth.  
```{r Test3, include=T, eval=T, echo=T}
#   The Combined Effects of Supplement & Dose on Tooth Growth
    t.test(ToothGrowth$len[ToothGrowth$supp=="OJ"], ToothGrowth$len[ToothGrowth$supp=="VC"],
           con.levle=0.95, paired=F);



```
####4. State your conclusions and the assumptions needed for your conclusions. 
  

