##Simulation Exercise
  
###Overview
The goal of this project is to investigate the exponential distribution in R and compare it with the Central Limit Theorem.  The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter.  The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda.  lambda = 0.2 for all of the simulations.  The distribution of averages of 40 exponentials will be investigated using 1,000 simulations.  
  
###Seed & Simulation Parameters
```{r include=T, eval=T, echo=T}
#   Seed & Simulation Parameters
    set.seed(5);
    sim = 1000;
    sample = 40;
    lambda = 0.2;
```
  
###1. Show the sample mean and compare it to the theoretical mean of the distribution.
  
The mean of the 1,000 simulated means is 5.043053, which is very close to theoretical mean of the distribution, which is 1/lambda = 1/0.20 = 5.  
```{r include=T, eval=T, echo=T}
#   Simulated Means
    simulatedMean <- replicate(sim, mean(rexp(sample, rate=lambda)));

#   Sample Mean
    sampleMean <- mean(simulatedMean);  sampleMean;
```

###2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
  
The sample variance is 0.6026047 is very similar to the theoretical distribution which is 0.625.  
```{r include=T, eval=T, echo=T}
#   Sample Variance  
    sampleVariance <- var(simulatedMean);  sampleVariance;

#   Theoretical Variance
    theoreticalVariance = (1/lambda)^2/sample;  theoreticalVariance;
```
  
###3. Show that the distribution is approximately normal.
  
The following histogram shows the distribution of means to be "approximately normal.""
```{r include=T, eval=T, echo=T}
#   Sample Distribution of Means
    hist(simulatedMean, prob=T, col="yellow", breaks=50, xlab="Sample mean", 
         main="Sample Mean Distribution");
    abline(v=0, col="red", lwd=4)
    lines(density(simulatedMean, adjust=2), col="blue", lwd=3)
    legend("topright", pch=12, col=c("red", "blue"), legend = c("Sample Mean Distribution",
        "Gaussian Approximation"))
```
  
```{r}
#   Sample Distribution
    hist(rexp(sample*sim,rate=lambda), prob=T, col="lightgreen", breaks=50, xlab="Sample", 
         main = "Sample Distribution")
    lines(density(rexp(sample*sim,rate=lambda), adjust=2), col="darkblue", lwd=3)
```

