##Simulation Exercise
  
###Overview
The goal of this project is to investigate the exponential distribution in R and compare it with the Central Limit Theorem.  The exponential distribution is simulated in R with rexp(n, lambda) where lambda is the rate parameter.  The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda; lambda = 0.20 for all of the simulations.  The distribution of means of 40 exponentials will be investigated using 1,000 simulations.  

```{r echo=F}
    library(ggplot2);
```
  
###Seed & Simulation Parameters
```{r include=T, eval=T, echo=T}
#   Seed & Simulation Parameters
    set.seed(777);
    sim = 1000;
    sample = 40;
    lambda = 0.20;
```
  
###1. Show the sample mean and compare it to the theoretical mean of the distribution.
  
The sample mean of the 1,000 simulated means is 4.969789.  This is extremely close to theoretical mean of the distribution, 5.  
```{r include=T, eval=T, echo=T}
#   Theoretical Mean
    theoreticalMean <- 1/lambda;  theoreticalMean;

#   Sample Mean
    simulatedMean <- replicate(sim, mean(rexp(sample, rate=lambda)));
    sampleMean <- mean(simulatedMean);  sampleMean;
```

###2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
  
The sample variance is 0.6384832.  This is fairly close to the theoretical distribution, which is 0.625.  
```{r include=T, eval=T, echo=T}
#   Sample Variance  
    sampleVariance <- var(simulatedMean);  sampleVariance;

#   Theoretical Variance
    theoreticalVariance = (1/lambda)^2/sample;  theoreticalVariance;
    
# Histogram: Sample vs Theoretical Mean
    hist(simulatedMean,  freq=T, main="Distribution of Sample Means", 
         xlab="Sample Means", col="lightblue");
    abline(v=sampleMean, col='blue', lwd=2, lty=1);
    abline(v=theoreticalMean, col='red', lwd=2, lty=4);
    legend('topright', c("Sample Mean", "Theoretical Mean"),
        col=c("blue", "red"), lwd=c(2, 2), lty=c(1, 4), bty="sim"); 
```
  
###3. Show that the distribution is approximately normal.
  
The following histogram shows the distribution of means to be "approximately normal."
```{r include=T, eval=T, echo=T}
#   Histogram: Sample Mean vs Theoretical Mean
    hist(simulatedMean,  prob=T, main="Distribution of Sample Means", 
         xlab="Sample Means", col="lightblue");
    lines(density(simulatedMean), lwd=2, lty=1, col="blue")
    legend('topright', c("Sample Mean", "Theoretical Mean"), col=c("blue", "red"),
        lwd=c(2, 2), lty=c(1, 4));
    xfit <- seq(min(simulatedMean), max(simulatedMean), length=100);
    yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(sample)));
    lines(xfit, yfit, pch=22, col="red", lwd=2, lty=4);
```


